Investigating Fire Incidents Severity Levels in Toronto: Enhancing Resource Allocation for Fire Response

### 2024-12-15 paper/paper.qmd
fix discussion

In fact, the cost of missing high-severity fires is far greater than the cost of false positives. While false positives may lead to overreaction, failing to detect a severe fire could escalate into a major disaster, resulting in significant financial losses and potentially causing more casualties. Therefore, a model that is tuned to minimize false negatives is preferable, as it prioritizes catching severe fires—even at the expense of some false positives. However, I will continue to fit additional models to strike a better balance between specificity and sensitivity.

@@

1.  **Classification Bias**: Classification bias occurs when events are mis-classified into categories that do not accurately represent the event itself [@brogden1950theory]. This leads to incorrect or misleading conclusions. For example, in the dataset, there is a variable called Possible_Cause, which has over 100 categories, including categories like "Undetermined" and "Suspected Vandalism." When categorizing causes into these categories, ambiguity arises. There may not be enough evidence to definitively classify a cause, and this uncertainty can result in bias. If we were to generalize these 100 causes into fewer categories for prediction purposes, one of the generalized categories could be biased, misrepresenting the actual distribution of causes. This is one of the reasons why I decided to remove this variable from the dataset, besides the fact that it contains many missing values (NAs).
2.  **Selective Reporting Bias**: Reporting bias occurs when data does not accurately reflect reality due to incomplete reporting [@saini2014selective]. In our case, this is particularly relevant for civilian and firefighter casualties related to different fire incidents. It's possible that some fire incidents are not reported, especially smaller ones that don't require significant resources. On the other hand, large, tragic fire incidents are often over-reported or emphasized more than necessary. As a result, the reported number of casualties may be skewed. These discrepancies in the data can affect the definition of what is considered a "severe" fire. If we use these skewed measurements to make predictions or draw conclusions, the resulting insights may be biased.

>>>>

In fact, the cost of missing high-severity fires is far greater than the cost of false positives. While false positives may lead to overreaction, failing to detect a severe fire could escalate into a major disaster, resulting in significant financial losses and potentially causing more casualties. Therefore, a model that is tuned to minimize false negatives is preferable, as it prioritizes catching severe fires even at the expense of some false positives. However, I will continue to fit additional models to strike a better balance between specificity and sensitivity.

@@

1.  **Classification Bias**: Classification bias occurs when events are misclassified into categories that do not accurately represent the event itself [@brogden1950theory]. This leads to incorrect or misleading conclusions. For example, in the dataset, there is a variable called Possible_Cause, which has over 100 categories, including categories like "Undetermined" and "Suspected Vandalism." When categorizing causes into these categories, ambiguity arises. There may not be enough evidence to definitively classify a cause, and this uncertainty can result in bias. If we were to generalize these 100 causes into fewer categories for prediction purposes, one of the generalized categories could be biased, misrepresenting the actual distribution of causes. This is one of the reasons why I decided to remove this variable from the dataset, besides the fact that it contains many missing values (NAs).
2.  **Selective Reporting Bias**: Reporting bias occurs when data does not accurately reflect reality due to incomplete reporting [@saini2014selective]. In our case, this is particularly relevant for civilian and firefighter casualties related to different fire incidents. It's possible that some fire incidents are not reported, especially smaller ones that don't require significant resources. On the other hand, large, tragic fire incidents are often over-reported or emphasized more than necessary. As a result, the reported number of casualties may be skewed. These discrepancies in the data can affect the definition of what is considered a "severe" fire. If we use these skewed measurements to make predictions or draw conclusions, the resulting insights may be biased.
